---
title: "Vertical niches 2.0"
author: "Brunno F Oliveira^1^; Brett R Scheffers^2^;"
date: "May 9, 2017 **1** Department of Wildlife Ecology and Conservation, University of Florida/IFAS, Gainesville, FL 32611, USA. **Corresponding author:** (brunno.oliveira@me.com)"
output:
  pdf_document: default
theme: united
toc: yes
---

*** 

\newpage



# Packages versions:
```{r info,message=FALSE, echo=F}
info <- sessionInfo()
```

We used `r info[1]$R$ver` and the following packages:

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
rm(list=ls())
gc()

list.of.packages <- c("picante","ape","raster","maptools","rgeos","stringr","rgdal","ggplot2","plyr","grid","gridExtra","multcompView","caper","geiger","phytools","knitr","maps","parallel","doParallel","apTreeshape","data.table","ppcor","hier.part", "relaimpo","MASS","bestglm")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

sapply(list.of.packages, require, character.only = TRUE)
```

# set directory
```{r "setup", include=FALSE}

savedatafolder <- "C:/Users/boliveir/Dropbox (UFL)/Arboreality_II/Data/" # address to folder to save results
externHD <- 'G:/GIS/' # address to external HD
opts_knit$set(root.dir = "C:/Users/boliveir/Dropbox (UFL)/Arboreality_II")
setwd("C:/Users/boliveir/Dropbox (UFL)/Arboreality_II")

```

# RData
```{r "setup", include=FALSE}
load("Analyses_data.RData")
```

# Load functions
```{r "setup", include=FALSE}
mundi <- readShapePoly(paste(externHD,"/Shp files/Mundi_contour.shp",sep=""))
#mundi <- map("world", interior=F, fill=F, ylim=c(-60, 90), mar=c(0,0,0,0), plot = F)

grid_arrange_shared_legend <- function(..., ncol = length(list(...)), nrow = 1, position = c("bottom", "right")) {

  plots <- list(...)
  position <- match.arg(position)
  g <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
  lheight <- sum(legend$height)
  lwidth <- sum(legend$width)
  gl <- lapply(plots, function(x) x + theme(legend.position="none"))
  gl <- c(gl, ncol = ncol, nrow = nrow)

  combined <- switch(position,
                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
                                            legend,
                                            ncol = 1,
                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
                                           legend,
                                           ncol = 2,
                                           widths = unit.c(unit(1, "npc") - lwidth, lwidth)))

  grid.newpage()
  grid.draw(combined)

  # return gtable invisibly
  invisible(combined)

}
```

# Load community data
```{r com_data, eval=T, message=TRUE}
# Birds
birds <- read.csv("Data/birds_comvars.csv")[,-1]
birds.a <- data.frame(birds[,c(1:2)],
                      log(birds[,c(3:5)]+1), # WTemp > range
                      birds[,6], # ses.vert
                      log(birds[,c(7:9)]+1), #arb/ter/fos
                      birds[,c(10:13)], # Parb/Pter/Pfos+mean.temp
                      log(birds[,c(14:21)]+1), # clim vars + rich
                      birds[,c(22:23)]) # realm & bioream
birds.a <- data.frame(birds[,c(1:2)],
                      scale(birds[,c(3:21)]),
                      birds[,c(22:23)]) # realm & bioream
names(birds.a) <- names(birds)


# Amphibians
amphibians <- read.csv("Data/amphibians_comvars.csv")[,-1]
amphibians <- amphibians[which(amphibians$Rich>5),]
amphibians$ses.vert <- amphibians$ses.vert.realm
amphibians$arb <- amphibians$P.arb * amphibians$Rich
amphibians$ter <- amphibians$P.ter * amphibians$Rich
amphibians$fos <- amphibians$P.fos * amphibians$Rich
amphibians <- amphibians[,which(names(amphibians)%in%names(birds))]
columns <- names(amphibians)[as.numeric(na.omit(match(names(birds),names(amphibians))))]
amphibians <- amphibians[,columns]
amphibians.a <- data.frame(amphibians[,c(1:2)],
                           log(amphibians[,c(3:5)]+1), # WTemp > range
                           amphibians[,6], # ses.vert
                           log(amphibians[,c(7:9)]+1), #arb/ter/fos
                           amphibians[,c(10:13)], # Parb/Pter/Pfos+mean.temp
                           log(amphibians[,c(14:21)]+1), # clim vars + rich
                           amphibians[,c(22:23)]) # realm & bioream
amphibians.a <- data.frame(amphibians[,c(1:2)],
                           scale(amphibians[,c(3:21)]),
                           amphibians[,c(22:23)]) # realm & bioream
names(amphibians.a) <- names(amphibians)


# Mammals
mammals <- read.csv("Data/mammals_comvars.csv")[,-1]
mammals.a <- data.frame(mammals[,c(1:2)],
                        log(mammals[,c(3:5)]+1), # WTemp > range
                        mammals[,6], # ses.vert
                        log(mammals[,c(7:9)]+1), #arb/ter/fos
                        mammals[,c(10:13)], # Parb/Pter/Pfos+mean.temp
                        log(mammals[,c(14:21)]+1), # clim vars + rich
                        mammals[,c(22:23)]) # realm & bioream
mammals.a <- data.frame(mammals[,c(1:2)],
                        scale(mammals[,c(3:21)]),
                        mammals[,c(22:23)]) # realm & bioream
names(mammals.a) <- names(mammals)


# Reptiles
reptiles <- read.csv("Data/reptiles_comvars.csv")[,-1]
reptiles.a <- data.frame(reptiles[,c(1:2)],
                         log(reptiles[,c(3:5)]+1), # WTemp > range
                         reptiles[,6], # ses.vert
                         log(reptiles[,c(7:9)]+1), #arb/ter/fos
                         reptiles[,c(10:13)], # Parb/Pter/Pfos+mean.temp
                         log(reptiles[,c(14:21)]+1), # clim vars + rich
                         reptiles[,c(22:23)]) # realm & bioream
reptiles.a <- data.frame(reptiles[,c(1:2)],
                         scale(reptiles[,c(3:21)]),
                         reptiles[,c(22:23)]) # realm & bioream
names(reptiles.a) <- names(reptiles)

```

# Load species data
```{r com_data, eval=T, message=TRUE}
# Birds
birds.sp <- read.csv("Data/birds_spvars.csv")[,-1]

# Amphibians
amphibians.sp <- read.csv("Data/amphibians_spvars.csv")[,-1]

# Mammals
mammals.sp <- read.csv("Data/mammals_spvars.csv")[,-1]

# Reptiles
reptiles.sp <- read.csv("Data/reptiles_spvars.csv")[,-1]
```

# Load realm & biorealm data
```{r com_data, eval=T, message=TRUE}
# Realm
# Birds
birds.realm <- read.csv("Data/birds_Realmvars.csv")[,-1]

# Amphibians
amphibians.realm <- read.csv("Data/amphibians_Realmvars.csv")[,-1]

# Mammals
mammals.realm <- read.csv("Data/mammals_Realmvars.csv")[,-1]

# Reptiles
reptiles.realm <- read.csv("Data/reptiles_Realmvars.csv")[,-1]

########
# biorealm
# Birds
birds.biorealm <- read.csv("Data/birds_biorealmvars.csv")[,-1]

# Amphibians
amphibians.biorealm <- read.csv("Data/amphibians_biorealmvars.csv")[,-1]

# Mammals
mammals.biorealm <- read.csv("Data/mammals_biorealmvars.csv")[,-1]

# Reptiles
reptiles.biorealm <- read.csv("Data/reptiles_biorealmvars.csv")[,-1]
```

# Stack Maps
```{r, message=FALSE, warning=FALSE, include=FALSE}
# Get XY in Mollweide equal-area projection
mapa <- raster(paste(sep="",externHD,"/Mollweide/mundo_1Dg.grd")) 

s.birds <- stack()
for(n in 3:ncol(birds)){
  d <- data.frame(birds[,1:2],birds[,n])
  ras <- rasterFromXYZ(xyz=d,digits=10,crs=CRS("+proj=cea +datum=WGS84"))
  ras <- projectRaster(ras, mapa)
  extent(ras) <- extent(mapa)
  s.birds <- addLayer(s.birds,ras)
  rm(d,n,ras)
}
names(s.birds) <- names(birds)[-1:-2]

s.amphibians <- stack()
for(n in 3:ncol(amphibians)){
  d <- data.frame(amphibians[,1:2],amphibians[,n])
  ras <- rasterFromXYZ(xyz=d,digits=10,crs=CRS("+proj=cea +datum=WGS84"))
  ras <- projectRaster(ras, mapa)
  extent(ras) <- extent(mapa)
  s.amphibians <- addLayer(s.amphibians,ras)
  rm(d,n,ras)
}
names(s.amphibians) <- names(amphibians)[-1:-2]

s.mammals <- stack()
for(n in 3:ncol(mammals)){
  d <- data.frame(mammals[,1:2],mammals[,n])
  ras <- rasterFromXYZ(xyz=d,digits=10,crs=CRS("+proj=cea +datum=WGS84"))
  ras <- projectRaster(ras, mapa)
  extent(ras) <- extent(mapa)
  s.mammals <- addLayer(s.mammals,ras)
  rm(d,n,ras)
}
names(s.mammals) <- names(mammals)[-1:-2]

s.reptiles <- stack()
for(n in 3:ncol(reptiles)){
  d <- data.frame(reptiles[,1:2],reptiles[,n])
  ras <- rasterFromXYZ(xyz=d,digits=10,crs=CRS("+proj=cea +datum=WGS84"))
  ras <- projectRaster(ras, mapa)
  extent(ras) <- extent(mapa)
  s.reptiles <- addLayer(s.reptiles,ras)
  rm(d,n,ras)
}
names(s.reptiles) <- names(reptiles)[-1:-2]

```

# Plot maps verticality
```{r maps, echo=TRUE, fig.height=8, fig.width=10, message=FALSE, warning=F}
# Amphibians
amphibians_SR_classint <- BAMMtools::getJenksBreaks(amphibians$Rich, 20) # Get class intervals RICH
amphibians_SR_classint <- unique(amph_SR_classint) # avoid identical breaks
# birds
birds_SR_classint <- BAMMtools::getJenksBreaks(birds$Rich, 20) # Get class intervals RICH
birds_SR_classint <- unique(amph_SR_classint) # avoid identical breaks
# mammals
mammals_SR_classint <- BAMMtools::getJenksBreaks(mammals$Rich, 20) # Get class intervals RICH
mammals_SR_classint <- unique(amph_SR_classint) # avoid identical breaks
# reptiles
reptiles_SR_classint <- BAMMtools::getJenksBreaks(reptiles$Rich, 20) # Get class intervals RICH
reptiles_SR_classint <- unique(amph_SR_classint) # avoid identical breaks

par(mfrow=c(3,2),mar=c(0,0,1,0))

# Amphibians
{plot(s.amphibians$Rich, main='Amphibian Richness',axes=F,box=F, col=colorRampPalette(c('blue','lightblue','yellow','red'))(20))
map(mundi,add=T,cex=.5)}
{plot(s.amphibians$ses.vert, main='Amphibian Verticality',axes=F,box=F, col=colorRampPalette(c('red','blue','green'))(20))
map(mundi,add=T,cex=.5)}

# Birds
{plot(s.birds$Rich, main='Birds Richness',axes=F,box=F, col=colorRampPalette(c('blue','lightblue','yellow','red'))(20))
map(mundi,add=T,cex=.5)}
{plot(s.birds$ses.vert, main='Birds Verticality',axes=F,box=F, col=colorRampPalette(c('red','blue','green'))(20))
map(mundi,add=T,cex=.5)}

# Mammals
{plot(s.mammals$Rich, main='Non-volant Mammals Richness',axes=F,box=F, col=colorRampPalette(c('blue','lightblue','yellow','red'))(20))
map(mundi,add=T,cex=.5)}
{plot(s.mammals$ses.vert, main='Non-volant Mammals Verticality',axes=F,box=F, col=colorRampPalette(c('red','blue','green'))(20))
map(mundi,add=T,cex=.5)}

# Reptiles
{plot(s.reptiles$Rich, main='Reptiles Richness',axes=F,box=F, col=colorRampPalette(c('blue','lightblue','yellow','red'))(20))
map(mundi,add=T,cex=.5)}
{plot(s.reptiles$ses.vert, main='Reptiles Verticality',axes=F,box=F, col=colorRampPalette(c('red','blue','green'))(20))
map(mundi,add=T,cex=.5)}
```

## Inset latitude
```{r latitude, echo=FALSE, fig.height=6, fig.width=3, message=FALSE, warning=FALSE}
# Amphibians
tempdata <- aggregate(amphibians[,c("P.arb","P.fos","P.ter")],
                      list(amphibians$y), mean)
tempdata2 <- data.frame(lat=rep(tempdata$Group.1,3),
                        prop=c(tempdata$P.arb,tempdata$P.ter,tempdata$P.fos),
                        vert=c(rep("arb",nrow(tempdata)),
                               rep("ter",nrow(tempdata)),
                               rep("fos",nrow(tempdata))))

histarbfos.a <-
  ggplot(tempdata2, aes(lat, prop, colour = vert)) +
  #geom_line()+
  geom_smooth(span=.3, size = .5)+
  scale_colour_manual(values = c("green","red","blue"))+
  theme_classic()+coord_flip()+
  labs(x = "", y = "")+
  theme(legend.position="none")

tempdata <- aggregate(amphibians$ses.vert, list(amphibians$y), function(x) mean(x, na.rm=TRUE))
histvert.a <-
  ggplot(tempdata, aes(Group.1, x)) +
  ggtitle(" ") +
  geom_hline(yintercept=0,linetype = "dashed") + 
  geom_smooth(span=.3, colour = 'black', size = .5)+
  theme_classic()+coord_flip()+
  labs(y = "", x = "")

# Birds
tempdata <- aggregate(birds[,c("P.arb","P.fos","P.ter")],
                      list(birds$y), mean)
tempdata2 <- data.frame(lat=rep(tempdata$Group.1,3),
                        prop=c(tempdata$P.arb,tempdata$P.ter,tempdata$P.fos),
                        vert=c(rep("arb",nrow(tempdata)),
                               rep("ter",nrow(tempdata)),
                               rep("fos",nrow(tempdata))))

histarbfos.b <-
  ggplot(tempdata2, aes(lat, prop, colour = vert)) +
  #geom_line()+
  geom_smooth(span=.3, size = .5)+
  scale_colour_manual(values = c("green","red","blue"))+
  theme_classic()+coord_flip()+
  labs(x = "", y = "")+
  theme(legend.position="none")

tempdata <- aggregate(birds$ses.vert, list(birds$y), function(x) mean(x, na.rm=TRUE))
histvert.b <-
  ggplot(tempdata, aes(Group.1, x)) +
  ggtitle(" ") +
  geom_hline(yintercept=0,linetype = "dashed") + 
  geom_smooth(span=.3, colour = 'black', size = .5)+
  theme_classic()+coord_flip()+
  labs(y = "", x = "")

# Mammals
tempdata <- aggregate(mammals[,c("P.arb","P.fos","P.ter")],
                      list(mammals$y), mean)
tempdata2 <- data.frame(lat=rep(tempdata$Group.1,3),
                        prop=c(tempdata$P.arb,tempdata$P.ter,tempdata$P.fos),
                        vert=c(rep("arb",nrow(tempdata)),
                               rep("ter",nrow(tempdata)),
                               rep("fos",nrow(tempdata))))

histarbfos.m <-
  ggplot(tempdata2, aes(lat, prop, colour = vert)) +
  #geom_line()+
  geom_smooth(span=.3, size = .5)+
  scale_colour_manual(values = c("green","red","blue"))+
  theme_classic()+coord_flip()+
  labs(x = "", y = "")+
  theme(legend.position="none")

tempdata <- aggregate(mammals$ses.vert, list(mammals$y), function(x) mean(x, na.rm=TRUE))
histvert.m <-
  ggplot(tempdata, aes(Group.1, x)) +
  ggtitle(" ") +
  geom_hline(yintercept=0,linetype = "dashed") + 
  geom_smooth(span=.3, colour = 'black', size = .5)+
  theme_classic()+coord_flip()+
  labs(y = "", x = "")

# Reptiles
tempdata <- aggregate(reptiles[,c("P.arb","P.fos","P.ter")],
                      list(reptiles$y), mean)
tempdata2 <- data.frame(lat=rep(tempdata$Group.1,3),
                        prop=c(tempdata$P.arb,tempdata$P.ter,tempdata$P.fos),
                        vert=c(rep("arb",nrow(tempdata)),
                               rep("ter",nrow(tempdata)),
                               rep("fos",nrow(tempdata))))

histarbfos.r <-
  ggplot(tempdata2, aes(lat, prop, colour = vert)) +
  #geom_line()+
  geom_smooth(span=.3, size = .5)+
  scale_colour_manual(values = c("green","red","blue"))+
  theme_classic()+coord_flip()+
  labs(x = "", y = "")+
  theme(legend.position="none")

tempdata <- aggregate(reptiles$ses.vert, list(reptiles$y), function(x) mean(x, na.rm=TRUE))
histvert.r <-
  ggplot(tempdata, aes(Group.1, x)) +
  ggtitle(" ") +
  geom_hline(yintercept=0,linetype = "dashed") + 
  geom_smooth(span=.3, colour = 'black', size = .5)+
  theme_classic()+coord_flip()+
  labs(y = "", x = "")
```

# Other metrics
```{r other_metrics, echo=FALSE, fig.height=8, fig.width=10, message=FALSE, warning=F}
par(mfrow=c(2,2))
## Birds
#Vertical Plasticity
{plot(s.birds$ses.vert.plast,main="Birds vertical flexibility", axes=F,box=F, col=colorRampPalette(c('red','blue','green'))(20))
map(mundi,add=T,cex=.5)}
#Food Plasticity
{plot(s.birds$ses.food.plast,main="Birds food flexibility",axes=F,box=F, col=colorRampPalette(c('red','blue','green'))(20))
map(mundi,add=T,cex=.5)}

## mammals
#Vertical Plasticity
{plot(s.mammals$ses.vert.plast,main="Mammals vertical flexibility",axes=F,box=F, col=colorRampPalette(c('red','blue','green'))(20))
map(mundi,add=T,cex=.5)}
#Food Plasticity
{plot(s.mammals$ses.food.plast,main="Mammals food flexibility",axes=F,box=F, col=colorRampPalette(c('red','blue','green'))(20))
map(mundi,add=T,cex=.5)}

```

# Models
## For richness
```{r warning= F, message=FALSE}
## Amphibians
fit <- glm(Rich ~ ., data = na.omit(amphibians.a[,-1:-2]))
summary(fit)
step <- stepAIC(fit, direction="backward")
step$anova # display results 

## Birds
fit <- glm(Rich ~ ., data = birds.a[,-1:-2])
summary(fit)
step <- stepAIC(fit, direction="backward")
step$anova # display results 

## Mammals
fit <- glm(Rich ~ ., data = mammals.a[,-1:-2])
summary(fit)
step <- stepAIC(fit, direction="backward")
step$anova # display results 
```

## For verticality
```{r warning= F}
## Amphibians
fit <- glm(ses.vert ~ ., data = na.omit(amphibians.a[,-1:-2]))
summary(fit)
step <- stepAIC(fit, direction="backward")
step$anova # display results 

## Birds
fit <- glm(ses.vert ~ ., data = birds.a[,-1:-2])
summary(fit)
step <- stepAIC(fit, direction="backward")
step$anova # display results 

## Mammals
fit <- glm(ses.vert ~ ., data = mammals.a[,-1:-2])
summary(fit)
step <- stepAIC(fit, direction="backward")
step$anova # display results 
```



# Hierarchical Partitioning 
calc.relimp calculates several relative importance metrics for the linear model.   


## For richness
```{r h_partitioning_rich, warning= F}
## Amphibians
limod <- lm(Rich ~ ., amphibians.a[,-1:-2])
metrics <- calc.relimp(limod, type = "lmg", rela = T)
as.matrix(100*metrics$lmg[order(100*metrics$lmg, decreasing = T)])

## Birds
limod <- lm(Rich ~ ., birds.a[,-1:-2])
metrics <- calc.relimp(limod, type = "lmg", rela = T)
as.matrix(100*metrics$lmg[order(100*metrics$lmg, decreasing = T)])

## Mammals
limod <- lm(Rich ~ ., mammals.a[,-1:-2])
metrics <- calc.relimp(limod, type = "lmg", rela = T)
as.matrix(100*metrics$lmg[order(100*metrics$lmg, decreasing = T)])
```

## For verticality
```{r h_partitioning_vert, warning= F}
## Amphibians
limod <- lm(ses.vert ~ ., amphibians.a[,-1:-2])
metrics <- calc.relimp(limod, type = "lmg", rela = T)
as.matrix(100*metrics$lmg[order(100*metrics$lmg, decreasing = T)])

## Birds
limod <- lm(ses.vert ~ ., birds.a[,-1:-2])
metrics <- calc.relimp(limod, type = "lmg", rela = T)
as.matrix(100*metrics$lmg[order(100*metrics$lmg, decreasing = T)])

## Mammals
limod <- lm(ses.vert ~ ., mammals.a[,-1:-2])
metrics <- calc.relimp(limod, type = "lmg", rela = T)
as.matrix(100*metrics$lmg[order(100*metrics$lmg, decreasing = T)])
```


# Plot correlations
```{r warning= F, echo=FALSE, fig.height=5, fig.width=10}
a <- 
  ggplot(amphibians, aes(ses.vert, Rich, color = Realm))+
  ggtitle("Amphibians") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(x = "SES vert", y = "Species richness")+
  scale_y_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5, show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

b <- 
  ggplot(birds, aes(ses.vert, Rich, color = Realm))+
  ggtitle("Birds") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(x = "SES vert", y = "Species richness")+
  scale_y_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

m <- 
  ggplot(mammals, aes(ses.vert, Rich, color = Realm))+
  ggtitle("Mammals") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(x = "SES vert", y = "Species richness")+
  scale_y_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

r <- 
  ggplot(reptiles, aes(ses.vert, Rich, color = Realm))+
  ggtitle("reptiles") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(x = "SES vert", y = "Species richness")+
  scale_y_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

grid_arrange_shared_legend(a, b, m, r, nrow=1)
```

```{r warning= F, echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
a <- 
  ggplot(amphibians, aes(mean.prec+1, ses.vert, color = Realm))+
  ggtitle("Amphibians") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Precipitation")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5, show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

b <- 
  ggplot(birds, aes(mean.prec+1, ses.vert, color = Realm))+
  ggtitle("Birds") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Precipitation")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

m <- 
  ggplot(mammals, aes(mean.prec+1, ses.vert, color = Realm))+
  ggtitle("Mammals") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Precipitation")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

r <- 
  ggplot(reptiles, aes(mean.prec+1, ses.vert, color = Realm))+
  ggtitle("reptiles") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Precipitation")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

grid_arrange_shared_legend(a, b, m, r, nrow=1)
```


```{r warning= F, echo=FALSE, fig.height=5, fig.width=10}
a <- 
  ggplot(amphibians, aes(veg+1, ses.vert, color = Realm))+
  ggtitle("Amphibians") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Vegetation")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5, show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

b <- 
  ggplot(birds, aes(veg+1, ses.vert, color = Realm))+
  ggtitle("Birds") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Vegetation")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

m <- 
  ggplot(mammals, aes(veg+1, ses.vert, color = Realm))+
  ggtitle("Mammals") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Vegetation")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

m <- 
  ggplot(mammals, aes(veg+1, ses.vert, color = Realm))+
  ggtitle("Mammals") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Vegetation")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

r <- 
  ggplot(reptiles, aes(veg+1, ses.vert, color = Realm))+
  ggtitle("reptiles") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Vegetation")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

grid_arrange_shared_legend(a, b, m, r, nrow=1)
```

```{r warning= F, echo=FALSE, fig.height=5, fig.width=10}
a <- 
  ggplot(na.omit(amphibians), aes(WTemp.cell+1, ses.vert, color = Realm))+
  ggtitle("Amphibians") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Niche width (temperature)")+
  #scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5, show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

b <- 
  ggplot(na.omit(birds), aes(WTemp.cell+1, ses.vert, color = Realm))+
  ggtitle("Birds") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Niche width (temperature)")+
  #scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

m <- 
  ggplot(na.omit(mammals), aes(WTemp.cell+1, ses.vert, color = Realm))+
  ggtitle("Mammals") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Niche width (temperature)")+
  #scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

r <- 
  ggplot(na.omit(reptiles), aes(WTemp.cell+1, ses.vert, color = Realm))+
  ggtitle("reptiles") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Niche width (temperature)")+
  #scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

grid_arrange_shared_legend(a, b, m, r, nrow=1)
```


```{r warning= F, echo=FALSE, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
a <- 
  ggplot(na.omit(amphibians), aes(range.size.cell+1, ses.vert, color = Realm))+
  ggtitle("Amphibians") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Range size")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5, show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

b <- 
  ggplot(na.omit(birds), aes(range.size.cell+1, ses.vert, color = Realm))+
  ggtitle("Birds") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Range size")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

m <- 
  ggplot(na.omit(mammals), aes(range.size.cell+1, ses.vert, color = Realm))+
  ggtitle("Mammals") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Range size")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

r <- 
  ggplot(na.omit(reptiles), aes(range.size.cell+1, ses.vert, color = Realm))+
  ggtitle("reptiles") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES vert", x = "Range size")+
  scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

grid_arrange_shared_legend(a, b, m, r, nrow=1)
```

```{r warning= F, echo=FALSE, fig.height=5, fig.width=10}
b <- 
  ggplot(na.omit(birds), aes(ses.vert, ses.food.plast, color = Realm))+
  ggtitle("Birds") + 
  geom_point(alpha = 0.1, legend = FALSE)+ 
  theme_classic()+
  labs(y = "SES food plasticity", x = "SES vert")+
  #scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "none",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

m <- 
  ggplot(na.omit(mammals), aes(ses.vert, ses.food.plast, color = Realm))+
  ggtitle("Mammals") + 
  geom_point(alpha = 0.1, legend = TRUE)+ 
  theme_classic()+
  labs(y = "SES food plasticity", x = "SES vert")+
  #scale_x_log10()+
  stat_smooth(aes(group=Realm, fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F)  +
  stat_smooth(color="black",method="lm",formula=y~x,level=0.95,se=T,size=1,show.legend=F)  +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2)))

grid_arrange_shared_legend(b, m, nrow=1)
```


# SAVE
```{r "setup", include=FALSE}
save.image("Analyses_data.RData")
```
