---
title: "Vertical niches 2.0 (reptiles)"
author: "Brunno F Oliveira^1^; Brett R Scheffers^2^;"
date: "May 9, 2017 **1** Department of Wildlife Ecology and Conservation, University of Florida/IFAS, Gainesville, FL 32611, USA. **Corresponding author:** (brunno.oliveira@me.com)"
output:
  pdf_document: default
theme: united
toc: yes
---

*** 

\newpage



# Packages versions:
```{r info,message=FALSE, echo=F}
info <- sessionInfo()
```

We used `r info[1]$R$ver` and the following packages:

```{r packages, echo=FALSE, message=FALSE, warning=FALSE, include=T}
rm(list=ls())
gc()

list.of.packages <- c("picante","ape","raster","maptools","rgeos","stringr","rgdal","ggplot2","plyr","grid","gridExtra","multcompView","caper","geiger","phytools","knitr","maps","parallel","doParallel","apTreeshape","data.table","ppcor","hier.part", "relaimpo")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

sapply(list.of.packages, require, character.only = TRUE)
```

#WHERE ARE YOU?
```{r "setup"}
# 1 brunno linux
# 2 scheffer's ws

i.am.at <- 1

if(i.am.at==1){
  savedatafolder <- "C:/Users/boliveir/Dropbox (UFL)/Arboreality_II/Data/" # address to folder to save results
  externHD <- 'G:/GIS/' # address to external HD
  opts_knit$set(root.dir = "C:/Users/boliveir/Dropbox (UFL)/Arboreality_II")
}

if(i.am.at==2){
  savedatafolder <- "C:/Users/boliveir/Dropbox (UFL)/Arboreality_II/Data/" # address to folder to save results
  externHD <- 'G:/GIS/' # address to external HD
  opts_knit$set(root.dir = "C:/Users/brunno.oliveira/Dropbox (UFL)/Arboreality_II")
}

```

## Land data
```{r , echo=F}
mundi <- rgdal::readOGR(paste(externHD,
                              "Shp files/ne_50m_land/ne_50m_land_no_artic.shp", 
                              sep = ""))
crs(mundi) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
mundi <- spTransform(mundi, CRS("+proj=cea +datum=WGS84"))

# REFERENCE PROJECTION MAP
mapa <- raster(xmn = -20592508, xmx = 20588492, ymn = -5743602, ymx = 6573398,
              crs = CRS("+proj=cea +datum=WGS84"))
res(mapa) <- 111000
```

# Load species pres/abs:
## Species occurrences
```{r species occr, eval=T, message=TRUE}
# Are you running the code for the first time?
answer <- "no"

if(answer=="no"){
  load(paste(savedatafolder,"reptiles",'_data.RData',sep=""))
}
if(answer=="yes"){
  occr <- read.csv(paste(savedatafolder,"Reptiles",'_occr_equal_area_30land.csv',sep=""))
  
  spp <- colnames(occr)[-1:-2]
  spp <- data.frame(gsub("[.]","_",spp))
  
  sppiucn <- spp
}
```
The original species occurrence dataset has `r nrow(spp)` species

# Check
```{r check, eval=T, message=TRUE}
#Species richness
if(answer=="no"){
  {test <- data.frame(XY,rowSums(occr))
  colnames(test)[1:2] <- c("x","y")
  coordinates(test) <- ~x+y
  gridded(test) <- TRUE
  test <- raster(test)}
  {plot(main="Richness", test)
    map(mundi,add=T,cex=.5)}
}
if(answer=="yes"){
  {test <- data.frame(occr[,c(1:2)],rowSums(occr[,c(-1:-2)]))
  colnames(test)[1:2] <- c("x","y")
  coordinates(test) <- ~x+y
  gridded(test) <- TRUE
  test <- raster(test)}
  {plot(main="Richness", test)
    map(mundi,add=T,cex=.5)}
}
```

## Traits
Using traits of lizards
Shai Meiri. 2018. Traits of lizards of the world: Variation around a successful evolutionary design. GEB.https://doi.org/10.1111/geb.12773 
```{r open_traits, eval=F}
### OPEN LARGE DATASET
trait <- data.frame(read.csv(paste(sep="",savedatafolder,"Meiri_-_Lizard_data.csv"), header=TRUE))

trait$species <- gsub(" ", "_", trait$species)

#names(trait)

# Get traits of interest
trait <- data.frame(trait[,c(1,11,12,20,21)]) 
trait$body <- unlist(apply(trait[,c(2,3)], 1, function(x) 
  max(x, na.rm=T)))

#names(trait)
```

The original trait dataset has `r nrow(trait)` species.  
These are the vertical layers in Meiri's traits:  
`unique(trait$substrate)`  

## Remove marice and semi-aquatic species
```{r traits, eval=F}
unique(trait$substrate)

# Remove: Aerial and Marine species
trait <- trait[-which(trait$substrate=="marine"),]
trait <- trait[-which(trait$substrate=="Semi_Aquatic"),]
```

# Verticality index
The verticality index represents the mean vertical layers occupied by species. Higher values represent species that forrage mainly in higher forest strata, whereas lower values represent species that forrage close to the ground.  

```{r}
# Calculate verticality index
vert.temp <- data.frame(vert = rep(NA, nrow(trait)))

vert.temp$vert[grep("Saxicolous",trait$substrate)] <- 2
vert.temp$vert[grep("Terrestrial",trait$substrate)] <- 2
vert.temp$vert[grep("Arboreal",trait$substrate)] <- 3
vert.temp$vert[grep("Arboreal&Saxicolous&Terrestrial",trait$substrate)] <- 3
vert.temp$vert[grep("Arboreal&Saxicolous",trait$substrate)] <- 3
vert.temp$vert[grep("Arboreal&Terrestrial",trait$substrate)] <- 2.5
vert.temp$vert[grep("Fossorial",trait$substrate)] <- 1
vert.temp$vert[grep("Cryptic",trait$substrate)] <- 2
vert.temp$vert[grep("Fossorial&Terrestrial",trait$substrate)] <- 1.5
vert.temp$vert[grep("Saxicolous&Terrestrial",trait$substrate)] <- 2
vert.temp$vert[grep("Fossorial&Saxicolous&Terrestrial",trait$substrate)] <- 1.5
vert.temp$vert[grep("Cryptic&Fossorial",trait$substrate)] <- 1
vert.temp$vert[grep("Cryptic&Terrestrial",trait$substrate)] <- 2

# fix Cryptic
# check if there is any information in the comments
# when Cryptic comes alone keep it as terrestrial
vert.temp <- data.frame(vert.temp,data.frame(trait))
cryptics <- which(vert.temp$substrate=="Cryptic")

for(i in cryptics){
  if(grepl("Terrestrial", vert.temp$substrate..comments[i])){
    vert.temp$vert[i] <- 2
  }
  if(grepl("Fossorial", vert.temp$substrate..comments[i])){
    vert.temp$vert[i] <- 1
  }
  if(grepl("Fossorial", vert.temp$substrate..comments[i])){
    vert.temp$vert[i] <- 1
  }
}

hist(vert)

trait <- data.frame(trait, vert=vert.temp$vert)
```


# Imputation
# for species without data on body size or arboreality, give the mean value across counterparts at the genus level.
```{r species ranges, eval=F, warning= F}
rownames(trait) <- NULL

toimputbody <- which(is.na(trait$body))

if(length(toimputbody)>0){
  allgenus <- rownames(trait)
  allgenus <- strsplit(allgenus,"_")
  allgenus <- unlist(lapply(allgenus, function(x) x[1]))

  for(i in toimputbody){
    traitsub <- which(allgenus==allgenus[i])
    trait$body[i] <- mean(trait$body[traitsub], na.rm = T)
  }
}

toimputvert <- which(is.na(trait$vert))
View(trait[toimputvert,])

if(length(toimputvert)>0){
  allgenus <- trait$species
  allgenus <- strsplit(allgenus,"_")
  allgenus <- unlist(lapply(allgenus, function(x) x[1]))

  for(i in toimputvert){
    traitsub <- which(allgenus==allgenus[i])
    trait$vert[i] <- mean(trait$vert[traitsub], na.rm = T)
  }
}

trait <- data.frame(species = trait$species,
                    body = trait$body,
                    vert = trait$vert)
trait <- na.omit(trait)

trait_names <- data.frame(trait$species)
```

## Create a list of species belonging to the two datasets (traits and species)
```{r eval=F}
names(spp) <- names(trait_names) <- "sps"

lista <- merge(data.frame(spp),data.frame(trait_names),by="sps")
lista <-as.vector(lista[,1])
lista <- unique(lista)
```

There are `r length(lista)` species in common between these two datasets (`(length(lista)*100)/nrow(spp)`% of all species present in the IUCN dataset). We use this species' list to subset the original species and trait data. This dataset will be used in the following analyses.

## Subseting species and trait datasets
```{r eval=F}
# For trait data
rownames(trait) <- trait_names$sps
trait <- trait[lista,] # Subset the species belonging to both datasets
trait <- trait[,-1]

## For distribution data
XY<-occr[,1:2] # get coordinates
XY <- as.data.frame(XY)
names(XY) <-c("x","y")

occr <- as.data.frame(occr[,-1:-2]) # get sps occurrences

names(occr) <- spp$sps

#############################
# create raster with the same extent of `Richness_Raster` (~1 degree)
# this will be usefull when getting environmental data
r <- test
res(r)

#############################

occr <- subset(occr, select=lista) # Subset the species beloguing to both datasets

Rich <- rowSums(occr) # Get richness for every cell

gc() # free up more memmory
```


# Arb/ter/fos
```{r species ranges, eval=F, warning= F}
trait$arb <- 
```

# Species range size
```{r species ranges, eval=F, warning= F}
range.size <- colSums(occr)
```

# Load environmental variables
## Climate change velocity (Sandel et al. 2011)
Downloaded from [Dryad](http://datadryad.org/resource/doi:10.5061/dryad.b13j1).
Citation: Sandel, B. et al. (2011) The influence of Late Quaternary climate-change velocity on species endemism. Science.
```{r get_velocity, eval=F}
velo.r <- raster(paste(sep="",externHD,'#Environment/climate_velocity/Velocity.tif'))
velo.r <- projectRaster(velo.r, crs = crs(mapa, asText = T))

velo.r <- resample(velo.r, r, method= "bilinear")
velocity <- extract(velo.r,XY,method='simple')
```

## Current climate conditions (Hijmans et al. 2005) 
Downloaded from [Worldclim project](http://www.worldclim.org) at the resolution of 10 minutes of a degree ~ 1 degree_
Citation: Hijmans, R.J., et al. (2005) Very high resolution interpolated climate surfaces for global land areas. International Journal of Climatology
```{r get_current_climate, eval=F}
bio <- raster::getData("worldclim", var = "bio", res = 10)
# create raster with the same extent of bio (~1 degree)
r2 <- bio

bio <- projectRaster(bio, crs = crs(mapa, asText = T))

mean.temp.r <- resample(bio$bio1,r, method= "bilinear") # Annual mean temperature
max.temp.r <-  resample(bio$bio5,r, method= "bilinear") # Max Temperature of Warmest Month
min.temp.r <-  resample(bio$bio6,r, method= "bilinear") # Min Temperature of Warmest Month
sea.temp.r <- resample(bio$bio4,r, method= "bilinear") # temperature seasonality
diu.temp.r <- resample(bio$bio2,r, method= "bilinear") # diurnal temperature range

max.prec.r <- resample(bio$bio13,r, method= "bilinear") # Precipitation of Wettest Month
min.prec.r <- resample(bio$bio14,r, method= "bilinear") # Precipitation of Driest Month
mean.prec.r <- resample(bio$bio12,r, method= "bilinear") # Annual preciptation
sea.prec.r <- resample(bio$bio15,r, method= "bilinear") # Precipitation seasonality

mean.temp <- extract(mean.temp.r,XY,method='simple')
max.temp <- extract(max.temp.r,XY,method='simple')
min.temp <- extract(min.temp.r,XY,method='simple')
sea.temp <- extract(sea.temp.r,XY,method='simple')
diu.temp <- extract(diu.temp.r,XY,method='simple')

mean.prec <- extract(mean.prec.r,XY,method='simple')
min.prec <- extract(min.prec.r,XY,method='simple')
max.prec <- extract(max.prec.r,XY,method='simple')
sea.prec <- extract(sea.prec.r,XY,method='simple')
```


## Canopy height and biomass at 1km spatial resolution.
Downloaded from [NASA](http://lidarradar.jpl.nasa.gov/).  
Citation: Simard, M. et al. (2011), "Mapping forest canopy height globally with spaceborne lidar", Journal of Geophysical Research, VOL. 116, G04021. 
```{r get_tree_heigh, eval=F}
veg.r <- raster(paste(sep="",externHD,'#Environment/NASA veg height/asc/veg-height_1dg-equal.asc'))
veg.r <- resample(veg.r, bio, method= "bilinear")

veg <- extract(veg.r,XY,method='simple')

tden.r <- raster(paste(sep="",externHD,'#Environment/Tree_density_Crowther_Nature/veg-density_1dg-equal.asc'))
tden.r <- resample(tden.r, bio, method= "bilinear")

tden <- extract(tden.r,XY,method='simple')


```


## Evelation data at 1 degree.
Downloaded from [NASA](https://asterweb.jpl.nasa.gov/gdem.asp).
```{r get_elevation, eval=F}
elev.r <- raster(paste(sep="",externHD,'Elevation/GDEM-10km-BW.tif'))
elev.r <- projectRaster(elev.r, crs = crs(mapa, asText = T))
elev.r <- resample(elev.r, bio, method= "bilinear")

elev <- extract(elev.r,XY,method='simple')
```


## Realms
```{r get_realm, warning= F, eval=F}
Reanames <- readShapePoly("D:/GIS/CMEC regions & realms/newRealms.shp")
crs(Reanames) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
extent(mapa) <- extent(Reanames)
Reanames<-data.frame(Reanames$Realm)
names(Reanames)<-'Realm'
Reanames<-cbind(Realms=rownames(Reanames),Reanames)
#2) Extract values
Realm <- readShapePoly("D:/GIS/CMEC regions & realms/newRealms.shp")
crs(Realm) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
realmnames<-sort(Realm$fullupgmar) # put alphabetic order because rasterize will put number in acordance with alphabetic order
extent(mapa) <- extent(Realm)
Realm <- rasterize(Realm,mapa,'fullupgmar')
Realm <- projectRaster(Realm, crs="+proj=cea +datum=WGS84")
Realm <- as.vector(raster::extract(Realm,XY,method='simple'))

#Some cells comprise more than one Realm and the extract function give a medium value for that cell. 
#This value have no meaning as we need the value of the Realm for each cell.
check.integer <- function(N){ # function for check if number is integer
!length(grep("[^[:digit:]]", as.character(N)))
}

for (i in seq_along(Realm)){ # put NA in non-integer numbers
if (check.integer(Realm[i]))
{Realm[i]<-Realm[i]
}else{Realm[i]<-NA}
}

#Dont know why...but some codes come with whitespaces -To see use: unique(Realm)
trim <- function( x ) { #Function to remove whitespaces
gsub("(^[[:space:]]+|[[:space:]]+$)", "", x)
}
Realm<-trim(Realm)

#Give names for Realm code...
Realm <- as.character(Reanames$Realm)[ match(Realm, c(1:11))]

### Solve problem with NAs between realms
Realm2 <- readShapePoly("D:/GIS/CMEC regions & realms/newRealms.shp")
crs(Realm2) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
Realm2 <- spTransform(Realm2, CRS("+proj=cea +datum=WGS84"))
XY <- SpatialPoints(XY)
crs(XY) <-"+proj=cea +datum=WGS84"
Realm2<-sp::over(XY,Realm2)
Realm2<-as.vector(Realm2$Realm)

#Give names for NAs based on the previous 
Realm3<-NA
for(i in 1:length(Realm2)){
if(is.na(Realm[i])){
Realm3[i]<-Realm2[i]
}else{
Realm3[i]<-Realm[i]
} 
}

Realm<-Realm3
rm(Realm2,Realm3)

# VISUALIZE
d <- na.omit(data.frame(as.data.frame(XY),Realm))
d$Realm <- as.numeric(d$Realm )
d <- rasterFromXYZ(d)
plot(d)  
```

## WWF Biomes
```{r get_tree_density}
biome <- readShapePoly(paste(sep="",externHD,"WWF_biomas_ecoregions/wwf_terr_ecos.shp"))
crs(biome) <-"+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0 "
biome <- spTransform(biome, CRS("+proj=cea +datum=WGS84"))
biome<-sp::over(XY,biome)
biome<-as.vector(biome$BIOME)

biome<-gdata::trim(biome)
biome<-round(as.numeric(biome),0)
biome[which(biome>14)] <- NA

# biomes names
bionames <- c(
"Tropical & Subtropical Moist Broadleaf Forests",
"Tropical & Subtropical Dry Broadleaf Forests", 
"Tropical & Subtropical Coniferous Forests",
"Temperate Broadleaf & Mixed Forests",
"Temperate Conifer Forests",
"Boreal Forests/Taiga",
"Tropical & Subtropical Grasslands, Savannas & Shrublands",
"Temperate Grasslands, Savannas & Shrublands",
"Flooded Grasslands & Savannas",
"Montane Grasslands & Shrublands",
"Tundra",
"Mediterranean Forests, Woodlands & Scrub",
"Deserts & Xeric Shrublands",
"Mangroves")

# biorealms
biorealm <- sapply(biome, function(x) bionames[biome[x]])

biorealm <- paste(Realm, biorealm, sep = " - ")
  
# # VISUALIZE
# d <- na.omit(data.frame(as.data.frame(XY),biome))
# d$biome <- as.numeric(d$biome )
# d <- rasterFromXYZ(d)
# plot(d)  
```

# Species metrics
Extract environmental features for every species
P.S.: This dataset will be useful for evolutionary analysis 

mean.temp.sp / min.temp.sp / max.temp.sp / sea.temp.sp = mean/min/max/seasonality temperature on species occurrences.

mean.prec.sp / min.prec.sp / max.prec.sp / sea.prec.sp = mean/min/max/seasonality preciptation on species occurrences.

WTemp = Temperature niche width = maximum temperature of the warmest month (Bio5) - minimum temperature of the coldest month (Bio6).

velocity.sp = mean climate change velocity within species ranges.

veg.sp = mean forest tree heigh within species ranges.

elev.sp = mean elevation (topography) within species ranges.

```{r get_species_vars, warning= F, eval=F}
# min temperature within species range
min.temp.sp <- apply(occr, 2, function(x) mean(na.omit(min.temp[which(x==1)])))
# max temperature within species range
max.temp.sp <- apply(occr, 2, function(x) mean(na.omit(max.temp[which(x==1)])))
# min preciptation within species range
min.prec.sp <- apply(occr, 2, function(x) mean(na.omit(min.prec[which(x==1)])))
# max preciptation within species range
max.prec.sp <- apply(occr, 2, function(x) mean(na.omit(max.prec[which(x==1)])))
# max elevation within species range
max.elev.sp <- apply(occr, 2, function(x) max(na.omit(elev[which(x==1)])))
# min elevation within species range
min.elev.sp <- apply(occr, 2, function(x) min(na.omit(elev[which(x==1)])))

# mean temperature within species range
mean.temp.sp <- apply(occr, 2, function(x) mean(na.omit(mean.temp[which(x==1)])))
# mean seasonality in temperature within species range
sea.temp.sp <- apply(occr, 2, function(x) mean(na.omit(sea.temp[which(x==1)])))
# species` temperature niche width
WTemp.sp <- apply(occr, 2, function(x) max(na.omit(max.temp[which(x==1)])) - min(na.omit(min.temp[which(x==1)])))
# mean climate change velocity within species range
velocity.sp <- apply(occr, 2, function(x) mean(na.omit(velocity[which(x==1)])))
# mean tree height within species` range
veg.sp <- apply(occr, 2, function(x) mean(na.omit(veg[which(x==1)])))
# mean elevation within species range
elev.sp <- apply(occr, 2, function(x) mean(na.omit(elev[which(x==1)])))
# range in elevation within species range
elev.range.sp <- apply(occr, 2, function(x) max(na.omit(elev[which(x==1)])) - min(na.omit(elev[which(x==1)])))
# mean latitude within species range
lat.sp <- apply(occr, 2, function(x) mean(na.omit(as.data.frame(XY)[,2][which(x==1)])))

## Create a dataframe to store species variables
sp.vars <- data.frame(trait,mean.temp.sp,sea.temp.sp,
                      min.temp.sp,max.temp.sp,min.prec.sp,max.prec.sp,max.elev.sp,min.elev.sp,
                      WTemp.sp,velocity.sp,veg.sp,
                      range.size,elev.sp,elev.range.sp,lat.sp)

write.csv(sp.vars, paste(savedatafolder,"reptiles","_spvars.csv",sep=""))
```

# Community metrics
```{r community_metrics, eval=F, message=FALSE}

# Mean community thermal niche width
WTemp.cell <- apply(occr, 1, function(x) mean(WTemp.sp[which(x==1)]))
# Mean community body size
body.cell <- apply(occr, 1, function(x) mean(trait$body[which(x==1)]))
# Mean community range size
range.size.cell <- apply(occr, 1, function(x) mean(range.size[which(x==1)]))
# Mean community Verticality
vert.cell <- apply(occr, 1, function(x) mean(trait$vert[which(x==1)]))
# N arboreals in community
arb <- apply(occr, 1, function(x) 
  length(which(trait$vert[which(x==1)] >= 2.5)))
# N terrestrials in community
ter <- apply(occr, 1, function(x) 
  length(which(trait$vert[which(x==1)] > 1.5 & 
          trait$vert[which(x==1)] < 2.5)))
# N fossorials in community
fos <- apply(occr, 1, function(x) 
  length(which(trait$vert[which(x==1)] <= 1.5)))
# Proportion arboreals in community
P.arb <- arb/Rich
# Proportion terrestrials in community
P.ter <- ter/Rich
# Proportion fossorials in community
P.fos <- fos/Rich

# SES of metrics
# 1. randomly pick n species from the same REALM and calculate the null metric
# 2. repeat 100x
# 3. calculate the mean and sd null verticality
tmp <- proc.time()

NCell <- nrow(occr)

ses.vert <- NA

for (i in 1:NCell) {
  cat("\r", i, "of", NCell)
  
  cells.sp <- which(occr[i, ] == 1) # species that occur in the cell i
  
  if (is.na(Realm[i])) {
    ses.vert[i] <- NA
  }
  else {
    Realm.cells <-
      as.numeric(which(Realm == Realm[i])) # cells in the realm
    occ.temp <- occr[Realm.cells, ] # realm subset
    pool <- which(colSums(occ.temp) > 0) # species that occur in the realm
    
    null.vert <- replicate(100, mean(sample(trait$vert[pool], Rich[i], replace = F)))
    ses.vert[i] <- (vert.cell[i] - mean(null.vert)) / sd(null.vert)
  }
}
time.total <- proc.time() - tmp


# Create a community dataset
com.vars <- data.frame(XY, WTemp.cell, body.cell, range.size.cell, 
                       ses.vert, arb, fos, ter, P.arb, P.fos, P.ter,
                       mean.temp,sea.temp,diu.temp,mean.prec,sea.prec,
                       veg, tden ,elev, velocity, Rich, Realm, biome)

# remove cells with less than 5 species
com.vars <- com.vars[which(com.vars$Rich>=5),]
com.vars <- com.vars[which(com.vars$veg>=1),]
com.vars <- na.omit(com.vars)

write.csv(com.vars, paste(savedatafolder,"reptiles","_comvars.csv",sep=""))

names(com.vars)
```

# create occr by realm and biorealm
```{r Realm_occr}
# occr Realm
teste <- data.frame(occr,Realm)
teste <- aggregate(.~Realm, teste, sum)
teste2 <- data.frame(Realm=teste$Realm)
Realm2 <- cbind(teste2, ifelse(teste[,-1]>0, 1, 0))
Realm3 <- Realm2[,1]
teste3 <- Realm2[,-1]
rownames(teste3) <- Realm3

# Does at least one community has zero species?
any(apply(teste3, 1, sum)==0) 

occr_Realm <- as.data.frame(teste3) ; rm(teste3, teste2, teste)

# occr biobiorealm
teste <- data.frame(occr,biorealm)
teste <- aggregate(.~biorealm, teste, sum)
teste2 <- data.frame(biorealm=teste$biorealm)
biorealm2 <- cbind(teste2, ifelse(teste[,-1]>0, 1, 0))
biorealm3 <- biorealm2[,1]
teste3 <- biorealm2[,-1]
rownames(teste3) <- biorealm3

# Does at least one community has zero species?
any(apply(teste3, 1, sum)==0) 

occr_biorealm <- as.data.frame(teste3) ; rm(teste3, teste2, teste)
```

## calculate P.arb/fos per realm and biorealm
```{r biome_vars}
# Realm
NCell <- nrow(occr_Realm)

rich.Realm<-P.arb.Realm<-P.fos.Realm<-P.ter.Realm<-
N.arb.Realm<-N.fos.Realm<-N.ter.Realm<-rep(NA,NCell)

for (i in 1:NCell){

cat("\r",i,"of", NCell)

cells.sp <- which(occr_Realm[i,]==1) # species that occur in the cell i

rich.Realm[i] <- length(cells.sp)

# proportion of arboreal and fossorials in the cell
arbo.temp <- cells.sp[which(vert[cells.sp]>2)]
P.arb.Realm[i] <- length(arbo.temp) / length(cells.sp) 
N.arb.Realm[i] <- length(arbo.temp)

foso.temp <- cells.sp[which(vert[cells.sp]<2)]
P.fos.Realm[i] <- length(foso.temp) / length(cells.sp) 
N.fos.Realm[i] <- length(foso.temp)

tero.temp <- cells.sp[which(vert[cells.sp]==2)]
P.ter.Realm[i] <- length(tero.temp) / length(cells.sp) 
N.ter.Realm[i] <- length(tero.temp)
}


Realm.vars <- data.frame(
rich.Realm,P.arb.Realm,P.fos.Realm,P.ter.Realm,
N.arb.Realm,N.fos.Realm,N.ter.Realm)

write.csv(Realm.vars, paste(savedatafolder,"reptiles_Realmvars.csv",sep = ""))

# biorealm
NCell <- nrow(occr_biorealm)

rich.biorealm<-P.arb.biorealm<-P.fos.biorealm<-P.ter.biorealm<-
N.arb.biorealm<-N.fos.biorealm<-N.ter.biorealm<-rep(NA,NCell)

for (i in 1:NCell){

cat("\r",i,"of", NCell)

cells.sp <- which(occr_biorealm[i,]==1) # species that occur in the cell i

rich.biorealm[i] <- length(cells.sp)

# proportion of arboreal and fossorials in the cell
arbo.temp <- cells.sp[which(vert[cells.sp]>2)]
P.arb.biorealm[i] <- length(arbo.temp) / length(cells.sp) 
N.arb.biorealm[i] <- length(arbo.temp)

foso.temp <- cells.sp[which(vert[cells.sp]<2)]
P.fos.biorealm[i] <- length(foso.temp) / length(cells.sp) 
N.fos.biorealm[i] <- length(foso.temp)

tero.temp <- cells.sp[which(vert[cells.sp]==2)]
P.ter.biorealm[i] <- length(tero.temp) / length(cells.sp) 
N.ter.biorealm[i] <- length(tero.temp)
}


biorealm.vars <- data.frame(
rich.biorealm,P.arb.biorealm,P.fos.biorealm,P.ter.biorealm,
N.arb.biorealm,N.fos.biorealm,N.ter.biorealm)

write.csv(biorealm.vars, paste(savedatafolder,"reptiles_biorealmvars.csv",sep = ""))
```

# Create Raster layer stack
```{r}
s <- stack()
for(n in 3:ncol(com.vars)){
  d <- data.frame(com.vars[,1:2],com.vars[,n])
  ras <- rasterFromXYZ(xyz=d,digits=10,crs=crs(mapa, asText = T))
  s <- addLayer(s,ras)
  rm(d,n,ras)
}

names(s) <- names(com.vars)[-1:-2]

```

# Maps
```{r maps, warning= F}
#Species richness
{plot(s$Rich)
map(mundi,add=T,cex=.5)}

#Body size
{plot(log(s$body.cell+1))
map(mundi,add=T,cex=.5)}

#Range size
{plot(s$range.size.cell)
map(mundi,add=T,cex=.5)}

#Verticality
{plot(s$ses.vert)
map(mundi,add=T,cex=.5)}

#Vertical Plasticity
{plot(s$ses.vert.plast)
map(mundi,add=T,cex=.5)}

#Food Plasticity
{plot(s$ses.food.plast)
map(mundi,add=T,cex=.5)}

#Temp. niche width
{plot(s$WTemp.cell)
map(mundi,add=T,cex=.5)}

#tree height
{plot(s$veg)
map(mundi,add=T,cex=.5)}

```

# Correlations
```{r warning= F}
# Pearson correlation between verticality and variables.
data.frame(cor(scale(na.omit(com.vars[,c(-1,-2,-18)]))))[,1:7]
```
Verticality seems to correlate strongly with low seasonality, high preciptation and taller trees (i.e., the tropics).
Verticality correlates negativelly with range size, suggesting that arboreal reptiles are generaly small and endemic.
Verticality also is associated with Richness.

# Hierarchical Partitioning 
calc.relimp calculates several relative importance metrics for the linear model.
```{r h_partitioning, warning= F}
newdata <- data.frame(scale(na.omit(com.vars[,c(-1,-2,-18)])))

# For Richness
limod <- lm(Rich ~ ., newdata)
summary(limod)

metrics <- calc.relimp(limod, type = "lmg", rela = T)
as.matrix(100*metrics$lmg[order(100*metrics$lmg, decreasing = T)])

# For verticality
limod <- lm(ses.vert ~ ., newdata)
summary(limod)

metrics <- calc.relimp(limod, type = "lmg", rela = T)
as.matrix(100*metrics$lmg[order(100*metrics$lmg, decreasing = T)])

# For vert plasticity
limod <- lm(ses.vert.plast ~ ., newdata)
summary(limod)

metrics <- calc.relimp(limod, type = "lmg", rela = T)
as.matrix(100*metrics$lmg[order(100*metrics$lmg, decreasing = T)])

```


# Plot correlations
Red line represents linear regression line   
Green line represents SLOSS line
```{r warning= F, fig.width=5, fig.height=5}
# preciptation leads to high arboreality?
# precipitation leads to higher productivity canopies and may allow species to find food at trees
{plot(log(com.vars$mean.prec+1), com.vars$ses.vert)
abline(lm(ses.vert ~ log(mean.prec+1), com.vars),col='red', lwd=2)
lines(loess.smooth(log(com.vars$mean.prec+1), com.vars$ses.vert), col = 'green',lwd=2)}
summary(lm(ses.vert ~ log(mean.prec+1), com.vars))

##############################################
# vegetation structure leads to high arboreality
{plot(log(com.vars$veg+1), com.vars$ses.vert)
abline(lm(ses.vert ~ veg, com.vars),col='red', lwd=2)
lines(loess.smooth(com.vars$veg, com.vars$ses.vert), col = 'green',lwd=2)}
summary(lm(ses.vert ~ veg, com.vars))

##############################################
# small body species concentrate in the tropics where there are more arboreals
# large body are migratory, which is more common in temperate zones
{plot(log(com.vars$body.cell+1), com.vars$ses.vert)
abline(lm(ses.vert ~ log(body.cell+1), com.vars),col='red', lwd=2)
lines(loess.smooth(log(com.vars$body.cell+1), com.vars$ses.vert), col = 'green',lwd=2)}
summary(lm(ses.vert ~ log(body.cell+1), com.vars))

##############################################
# large range are migratory and more common in temperate regions
{plot(log(com.vars$range.size.cell+1), com.vars$ses.vert)
abline(lm(ses.vert ~ log(range.size.cell+1), com.vars),col='red', lwd=2)
lines(loess.smooth(log(com.vars$range.size.cell+1), com.vars$ses.vert), col = 'green',lwd=2)}
summary(lm(ses.vert ~ log(range.size.cell+1), com.vars))

##############################################
# vertical plasticity is correlated with arboreality
{plot(com.vars$ses.vert.plast, com.vars$ses.vert)
abline(lm(ses.vert ~ ses.vert.plast, com.vars),col='red', lwd=2)
lines(loess.smooth(com.vars$ses.vert.plast, com.vars$ses.vert), col = 'green',lwd=2)}
summary(lm(ses.vert ~ ses.vert.plast, com.vars))

##############################################
# arboreality is correlated with food plasticity
{plot(com.vars$ses.food.plast, com.vars$ses.vert)
abline(lm(ses.vert ~ ses.food.plast, com.vars),col='red', lwd=2)
lines(loess.smooth(com.vars$ses.food.plast, com.vars$ses.vert), col = 'green',lwd=2)}
summary(lm(ses.vert ~ ses.food.plast, com.vars))

##############################################
# vertical plasticity is coorelated with food plasticity
{plot(com.vars$ses.food.plast, com.vars$ses.vert.plast)
abline(lm(ses.vert ~ ses.food.plast, com.vars),col='red', lwd=2)
lines(loess.smooth(com.vars$ses.food.plast, com.vars$ses.vert), col = 'green',lwd=2)}
summary(lm(ses.vert ~ ses.food.plast, com.vars))

##############################################
# arboreality is correlated with richness
{plot(com.vars$ses.vert,log(com.vars$Rich+1))
abline(lm(log(Rich+1)~ses.vert, com.vars),col='red', lwd=2)
lines(loess.smooth(com.vars$ses.vert,log(com.vars$Rich+1)), col = 'green',lwd=2)}
summary(lm(ses.vert ~ Rich, com.vars))

##############################################
# vertical plasticity is correlated with richness
{plot(com.vars$ses.vert.plast,log(com.vars$Rich+1))
abline(lm(log(Rich+1)~ses.vert.plast, com.vars),col='red', lwd=2)
lines(loess.smooth(com.vars$ses.vert.plast,log(com.vars$Rich+1)), col = 'green',lwd=2)}
summary(lm(ses.vert.plast ~ Rich, com.vars))

##############################################
# food plasticity is correlated with richness
{plot(com.vars$ses.food.plast,log(com.vars$Rich+1))
abline(lm(log(Rich+1)~ses.food.plast, com.vars),col='red', lwd=2)
lines(loess.smooth(com.vars$ses.food.plast,log(com.vars$Rich+1)), col = 'green',lwd=2)}
summary(lm(ses.food.plast ~ Rich, com.vars))

##############################################
# large range sizes leads to large vertical plasticity?
ggplot(com.vars, aes(ses.vert.plast,range.size.cell,color = Realm))+
  geom_point(size=3, alpha = 0.1, legend = FALSE)+ 
  theme_classic()+  labs(x = "Vertical plasticity", y = "range size")+
  stat_smooth(aes(group=Realm,fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F) +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        #axis.title = element_text(face="bold"),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1)+
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=4)))

# large vertical plasticity leads to large temperature niche width?
ggplot(com.vars, aes(ses.vert.plast,WTemp.cell,color = Realm))+
  geom_point(size=3, alpha = 0.1, legend = FALSE)+ 
  theme_classic()+  labs(x = "Vertical plasticity", y = "Temperature niche width")+
  stat_smooth(aes(group=Realm,fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F) +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        #axis.title = element_text(face="bold"),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1)+
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=4)))

# large vertical plasticity leads to large food plasticity?
ggplot(com.vars, aes(ses.vert.plast,ses.food.plast,color = Realm))+
  geom_point(size=3, alpha = 0.1, legend = FALSE)+ 
  theme_classic()+  labs(x = "Vertical plasticity", y = "Food plasticity")+
  stat_smooth(aes(group=Realm,fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F) +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        #axis.title = element_text(face="bold"),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1)+
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=4)))


# large temperature niche width leads to large food plasticity?
ggplot(com.vars, aes(WTemp.cell,ses.food.plast,color = Realm))+
  geom_point(size=3, alpha = 0.1, legend = FALSE)+ 
  theme_classic()+  labs(x = "Temperature niche width", y = "Food plasticity")+
  stat_smooth(aes(group=Realm,fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F) +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        #axis.title = element_text(face="bold"),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1)+
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=4)))

# food plasticity decreases with temperature?
ggplot(com.vars, aes(mean.temp,ses.food.plast,color = Realm))+
  geom_point(size=3, alpha = 0.1, legend = FALSE)+ 
  theme_classic()+  labs(x = "Temperature", y = "Food plasticity")+
  stat_smooth(aes(group=Realm,fill = Realm),method="lm",formula=y~x,level=0.95,se=T,size=.5,show.legend=F) +
  theme(text = element_text(size = 14),legend.title = element_blank(),
        #axis.title = element_text(face="bold"),
        axis.text.x=element_text(colour="black", size = 12), 
        axis.text.y=element_text(colour="black", size = 12),
        legend.position = "right",
        aspect.ratio=1)+
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=4)))

```

# SAVE
```{r echo=F, eval=FALSE}
save.image(paste(savedatafolder,"reptiles",'_data.RData',sep=""))
```